\chapter{Project Plan}

Below is a brief outline of how we intend to proceed with this project.

\section{Key Milestones}

\subsection{Background}

\textit{\textbf{Time allocated: 4 weeks}} \hfill \textbf{DONE}

The background research serves as an important basis for this project, especially given the highly theoretical nature of the work I'll be doing. The collation of information into a concise document that I can refer to, as well as a database of relevant knowledge will be a great help when introducing novel concepts into Pony.

\subsection{New Formal Model}

\textit{\textbf{Time allocated: 4 weeks}}

The next step will be to formalize a new model for Pony based on the previous work done. This will incorporate ideas from Sylvan's, George's and Paul's models, as well as a few concept not yet modelled, such as recovery. 

\subsection{Proof of Soundness}

\textit{\textbf{Time allocated: 3 weeks}}

After the formal model has been completed, the next step will be to establish a proof of soundness. If a proof of soundness cannot be established, we may need to revisit the model.

\subsection{Investigation of Generics}

\textit{\textbf{Time allocated: 2 weeks}}

We have two options here; we can choose to expand the ideas presented in Paul's thesis, or to investigate the Materials/Shapes model and how it would apply to Pony.

\subsection{Design Model for Generics}

\textit{\textbf{Time allocated: 4 weeks}}

Based on the investigation, we will decide how we shall approach designing a generic version of our model. Once the model is designed, we shall try to prove it sound.


\subsection{Evaluation and Write-Up}

\textit{\textbf{Time allocated: 2 weeks}}

While I will endeavour to be constantly working on my report throughout the duration of this project, I have allocated the remaining time to ensuring the report is fully polished and my evaluation has been completed.
